---
title: "A: Computing Effect Sizes"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(metafor)
```

# Exercise A

Use the `metadat::dat.curtis1998` dataset in metadat. 
This dataset contains results from 27 studies on the
effects of CO2 levels on wood plant mass. 
See `?metadat::dat.curtis1998` for details.
   
1. Make a dataset with **standardized mean differences** (`measure = "SMD"`)
   as the effect size. Use the **sample effect size estimate** to compute
   the sampling error variance.

```{r}
dat_g <- escalc(
  data = dat.curtis1998,
  measure = "SMD", 
  vtype = "LS",
  m1i = m1i, sd1i = sd1i, n1i = n1i, 
  m2i = m2i, sd2i = sd2i, n2i = n2i
)

dat_g
```

   A. Construct confidence intervals for each effect size using a normal distribution.
   
```{r}
dat_g |> 
  mutate(
    CI_low = yi - qnorm(.975) * sqrt(vi),
    CI_high = yi + qnorm(.975) * sqrt(vi)
  )
```
   
   B. Why might using the sample effect size estimate to compute sampling error
      variance be a problem?
      
  - The sampling error formula includes effect size. If this is estimated poorly,
    then the sampling error will also be estimated poorly, distorting the study
    weights.


2. Make a new dataset with **standardized mean differences**, but compute the
   sampling error variance for studies by using the **average effect size** 
   across the included studies. Construct confidence intervals for each effect 
   size using a normal distribution.
   
```{r}
dat_g <- escalc(
  data = dat.curtis1998,
  measure = "SMD", 
  vtype = "AV",
  m1i = m1i, sd1i = sd1i, n1i = n1i, 
  m2i = m2i, sd2i = sd2i, n2i = n2i
)

dat_g |> 
  mutate(
    CI_low = yi - qnorm(.975) * sqrt(vi),
    CI_high = yi + qnorm(.975) * sqrt(vi)
  )
```


Use the `metadat::dat.bornmann2007` dataset in metadat. 
This dataset contains results from 66 studies on gender differences in 
grant awards across different fields. 
See `?metadat::dat.bornmann2007` for details.

3. Make a dataset with **log risk ratios** as the effect size.

```{r}
dat_lrr <- escalc(
  data = dat.bornmann2007,
  measure = "RR",
  ai = waward, n1i = wtotal,
  ci = maward, n2i = mtotal
)

dat_lrr
```   

4. Make a dataset with **log odds ratios** as the effect size.

```{r}
dat_lor <- escalc(
  data = dat.bornmann2007,
  measure = "OR",
  ai = waward, n1i = wtotal,
  ci = maward, n2i = mtotal
)

dat_lor
```   


Use the `metadat::dat.cohen1981` dataset in metafor. 
This dataset contains results from 20 studies on correlations
between teacher ratings of student aptitude and student performance.
See `?metadat::dat.cohen1981` for details.

5. Make a dataset with **unbiased correlation** as the effect size.
   Use the average effect size to compute the sampling error variances.

```{r}
dat_ucor <- escalc(
  data = dat.cohen1981,
  measure = "UCOR",
  vtype = "AV",
  ri = ri,
  ni = ni
)

dat_ucor
```

6. Make a dataset with **z-transformed correlation** as the effect size.
   Compute 95% confidence intervals for each study using a normal
   distribution, then back-transform to the correlation metric.

```{r}
dat_zcor <- escalc(
  data = dat.cohen1981,
  measure = "ZCOR",
  ri = ri,
  ni = ni
)

dat_zcor |> 
  mutate(
    CI_low = metafor::transf.ztor(yi - qnorm(.975) * sqrt(vi)),
    CI_high = metafor::transf.ztor(yi + qnorm(.975) * sqrt(vi))
  )
```
