---
title: "C: Heterogeneity"
format: html
editor_options: 
  chunk_output_type: console
---

# Exercise C

Use the `metadat::dat.bangertdrowns2004` dataset in metafor. 
This dataset contains results from 48 studies on the 
effectiveness of writing to learn interventions on academic achievement.
See `?metadat::dat.bangertdrowns2004` for details.

The dataset already contains computed `yi` and `vi` values, so no need for `escalc()`.

Fit a random effects model using these data.
Use either the `"HSk"` or `"DL"` estimator for tau. 
Then, fit a second model using the `"REML"` estimator for tau.

```{r}
dat_g <- dat.bangertdrowns2004

mod_hsk <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(id, author, year),
  method = "HSk", test = "knha",
  data = dat_g
)

mod_dl <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(id, author, year),
  method = "DL", test = "knha",
  data = dat_g
)

mod_reml <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(id, author, year),
  method = "REML", test = "knha",
  data = dat_g
)
```

   A. Compare the estimate of tau for the two models.
   
```{r}
confint(mod_hsk)
confint(mod_dl)
confint(mod_reml)
```

  - The HSk and DL estimators both estimate tau to be about .21. The REML
    estimator estimates it to be slightly larger at .22. This is typical, 
    as the closed-form estimators (HSk and DL) are somewhat negatively biased.
   
   B. Interpret the results of the REML model, including the regression
      coefficients and tau.
      
```{r}
mod_reml
confint(mod_reml)
```

   - The average treatment effect is estimated to be *g* = .22 standard deviations
     improvement in reading performance due to the reading intervention. Tau is
     estimated to be .22, indicating that treatment effects are estiamted to
     vary around the mean on average by about .22.
      
   C. Compute a confidence interval for the mean and a prediction interval
      for individual true outcomes. Interpret these.
      
```{r}
predict(mod_reml)
```

  - The confidence interval for the mean ranges from .12 to .32. This indicates
    uncertainty in the *average* treatment effect, the range where we expect the
    true mean effect to be. This interval indicates we are confident that the
    average effect of the reading intervention is small and positive.
  - The prediction interval for individual studies ranges from -.24 to .68. This
    indicates the range where we expect true individual study effects to fall.
    This indicates that although we are confident the average intervention effect
    is small and positive, we still expect some settings (i.e., studies) where 
    the treatment is very beneficial (.68) and some settings where it might be
    somewhat harmful (-.24).
      
   D. Compute a confidence interval for tau and interpret it.
      Also interpret tau^2, I^2, and H^2.
      
```{r}
mod_reml
confint(mod_reml)
```

  - Tau is estimated to be .22 with confidence interval [.16, .39].
    This is the standard deviation of true study effects around the mean.
    These values indicate a moderate amount of heterogeneity, with the uncertainty
    interval ranging from small to very large amounts of heterogeneity.
  - Tau^2 is the estimated *variance* of true study effects. It is typically 
    easier to to interpret tau instead.
  - I^2 is the estimated percent of the total variability that is not accounted
    for by sampling error and that may be due to moderators. It is a *relative*
    measure of heterogeneity and so should be interpreted cautiously.
    We estimated that 58% of the observed variability in effect sizes across
    studies is not due to artefacts (and may be due to moderators).
  - H^2 is the ratio of total variability to sampling (error) variability.
    It can be interpreted as a sort of signal to noise ratio.
      
   E. Create a forest plot for these results.
      Be sure to include a prediction interval.
      
```{r}
forest(mod_reml, addpred = TRUE)
```

   F. Compute BLUPs from this model. 
      What do these values mean?   
      Which studies will be adjusted the most?
      
```{r}
blups_reml <- blup(mod_reml) |> 
  as.data.frame() |> 
  rownames_to_column("study")

dat_blup <- bind_rows(
  observed = dat_g |> mutate(
    pred = yi,
    se = sqrt(vi),
    pi.lb = yi - qnorm(.975) * se, 
    pi.ub = yi + qnorm(.975) * se,
    study = paste(id, author, year)
  ),
  blup = blups_reml,
  .id = "type"
)

ggplot(dat_blup) +
  aes(y = study, x = pred, xmin = pi.lb, xmax = pi.ub, shape = type, color = type) +
  geom_pointrange(position = position_dodge(width = .75)) +
  see::theme_modern() +
  see::scale_color_okabeito()
```

  - These are the shrunken effects for each study -- our best estimate of the
    true effect for each study after adjusting for likely sampling error.
    The studies that are adjusted most will be the studies with (1) lots of
    sampling error, and (2) the most extreme effect sizes.