---
title: "D: Moderator Analyses Solutions"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(metafor)
```


# Exercise D

Use the `metadat::dat.bangertdrowns2004` dataset in metafor. 
This dataset contains results from 48 studies on the 
effectiveness of writing to learn interventions on academic achievement.
See `?metadat::dat.bangertdrowns2004` for details.

The dataset already contains computed `yi` and `vi` values, so no need for `escalc()`.

1. Fit a mixed effects model using these data with `length` as a continuous predictor.
   Use the `"REML"` estimator for tau.

```{r}
dat_g <- dat.bangertdrowns2004 |>
  mutate(
    grade = factor(
      grade,
      levels = 1:4,
      labels = c("Elementary", "Middle", "High school", "College")
    )
  ) |>
  drop_na(grade, length)

mod_length <- rma(
  yi = yi ~ 1 + length,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = dat_g
)

mod_length
```

  A. Is length of treatment related to treatment effectiveness? How so?
  
  - For each additional week of treatment, the size of the treatment effect 
    is expected to increase slightly, by *d* = .01 [95% CI = -.00, .03].

  B. What does the intercept mean in this model?
  
  - The expected treatment effect for a treatment of 0 weeks is 
    *d* = .07 [95% CI = -.10, .24]. A treatment cannot be zero weeks, so this
    intercept value is not directly applicable to any study
      
  C. Interpret R^2, F (or QM), and QE for this model.
  
  - *R*^2^ = .05, indicating that 5% of the residual heterogeneity across studies
    can be accounted for by length of treatment. The *F* test has *p* = .056,
    indicating that a null model where length has no effect is only modestly
    compatible with the data. The *Q*~E~ test has *p* < .0001, indicating that
    a model with no residual heterogeneity remaining is poorly supported by
    the data. Together, these results indicate that length likely has a modest
    impact on treatment effectiveness, but there is substantial heterogeneity
    in effects that remain unaccounted for.
   
  D. Compute a confidence interval for the mean and a prediction interval
     for individual study true outcomes. Interpret these.
     
```{r}
newmods <- create_newmods(
  mod_length,
  data = data.frame(length = c(1, 4, 10, 15, 24))
)

predict_length <- predict(mod_length, newmods = newmods) |>
  clean_rma_predictions(data = newmods)

predict_length
```

  - For a 1-week treatment, the average treatment effect is essentially zero at
    *d* = .08. The confidence interval ranges from -.08 to .24, indicating that
    the mean treatment effect might range anywhere from negligibly harmful (-.08)
    to weakly beneficial (.24). The prediction inteval ranges from .37 to .54,
    indicating that any specific intervention might be moderately harmful (-.37)
    to highly beneficial (.54).
  - For a 24-week treatment, the average treatment effect is moderately positive
    at *d* = .43. The confidence interval ranges from .18 to .67, indicating that
    the mean treatment effect is clearly beneficial and might range anywhere 
    from somewhat beneficial (.19) to strongly beneficial (.67). The prediction 
    inteval ranges from -.06 to .91, indicating that any specific intervention 
    might range from having a negligible effect (-.06) to being 
    extremely beneficial (.92).
      
  E. Compute a confidence interval for tau and interpret it.

```{r}
confint(mod_length)
```

  - tau is estimated to be .21 with confidence interval ranging .14 to .36. 
    This indicates that treatment effects are clearly heterogeneous, even
    after accounting for treatment length, with heterogeneity ranging from small
    (.14) to substantial (.36).

  F. Create a forest plot for these results.
     What do the grey diamonds for each study mean here?
     
```{r}
forest(
  mod_length, 
  order = length, 
  addfit = TRUE, 
  ilab = na.omit(select(dat.bangertdrowns2004, length)),
  ilab.xpos = -2
)
text(
  c(-5.2, -2, 5.5), 48, 
  c("Study", "Length", "g [95% CI]"), 
  cex = .8
)
```

  - The gray diamonds indicate the predicted mean treatment effect and 95%
    confidence interval for each study.
    

2. Compare effectiveness for different grade levels.

  A. Convert `grade` to a factor variable and relabel the levels from `1:4`, to
     `c("Elementary", "Middle", "High school", "College")`.
     
  - See above.
     
  B. Fit a meta-regression model with `grade` as a predictor. 
     What are the estimated mean outcomes for each grade level?
     What is tau?
     
```{r}
mod_grade <- rma(
  yi = yi ~ 1 + grade,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = dat_g
)

newmods <- create_newmods(
  mod_grade,
  data = expand_grid(
    grade = levels(dat_g$grade)
  )
)

predict_grade <- predict(mod_grade, newmods = newmods) |>
  clean_rma_predictions(data = newmods) |> 
  arrange(grade)

predict_grade

confint(mod_grade)
```
  
  - Treatment has similar effects among Elementary, High School, and College
    students, with mean values ranging from *d* = .22 to .31 and lower 95% 
    confidence bounds ranging from .07 to .09.
  - tau was fixed to be equal across grade levels. It was estimated to be
    .22 [95% CI .14, .35], indicating clear evidence for heterogeneity of
    treatment effects within grade levels.
     
  C. Fit separate subgroup models for the Elementary and High School studies.
     How do results of these models differ from the results from the meta-regression?
      
```{r}
mod_elementary <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = filter(dat_g, grade == "Elementary")
)

mod_highschool <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = filter(dat_g, grade == "High school")
)

predict(mod_elementary)
confint(mod_elementary)

predict(mod_highschool)
confint(mod_highschool)
```

  - Estimated mean values for Elementary and High School students were similar
    in both models. The confidence interval for High School students was wider
    in the subgroup model compared to the meta-regression. This is because 
    tau was estimated to be much larger in the High School group (tau = .37)
    compared to the pooled tau estimate from the meta-regression (tau = .22).
    
  D. Fit a location-scale meta-regression model `grade` as a predictor. 
     How do results of these models differ from the results (B) and (C)?
    
```{r}
mod_gradeSL <- rma(
  yi = yi ~ 1 + grade,
  vi = vi,
  scale = ~ 1 + grade,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = dat_g
)

newmods <- create_newmods(
  mod_gradeSL,
  data = expand_grid(
    grade = levels(dat_g$grade)
  )
)

predict_gradeSL <- predict(mod_gradeSL, newmods = newmods) |>
  clean_rma_predictions(data = newmods) |> 
  arrange(grade)

predict_gradeSL

confint(mod_gradeSL)
```

  - Results are essentially identical to the subgroup models from (C). 
    This is because tau is estimated separately for each grade group.
    
3. Bonus: Does the impact of treatment length vary by grade level?

```{r}
mod_lengthGrade <- rma(
  yi = yi ~ 1 + length * grade,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = dat_g
)

newmods <- create_newmods(
  mod_lengthGrade,
  data = expand_grid(
    length = c(1, 4, 10, 15, 24),
    grade = levels(dat_g$grade)
  )
)

predict_lengthGrade <- predict(mod_lengthGrade, newmods = newmods) |>
  clean_rma_predictions(data = newmods) |> 
  arrange(grade, length)

predict_lengthGrade
```

  - Intervention effectiveness does not increase with length for Elementary
    students, but does increase strongly with length for High School and College
    students. Interventions appear to be slightly to moderately harmful for
    Middle School students, but there is a lot of uncertainty in this result.


4. Bonus: Automating subgroup modeling in (2).

```{r}
## Automating the subgroup modeling
mods_grade <- dat_g |>
  arrange(grade) |>
  group_by(grade) |>
  nest() |>
  mutate(
    meta = list(rma(
      yi = yi ~ 1,
      vi = vi,
      slab = paste(author, year),
      method = "REML", test = "knha",
      data = data
    ))
  ) |>
  mutate(
    meta = setNames(meta, grade),
    k = map(meta, \(ma) ma$k),
    predict = map(meta, \(ma) clean_rma_predictions(predict(ma))),
    tau = map(meta, \(ma) get_tau_ci(ma))
  ) |>
  ungroup() |>
  tidyr::unnest(cols = c(k, predict, tau))

mods_grade
```

