---
title: "F: Diagnostics"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(metafor)
```


# Exercise F

Use the `metadat::dat.bangertdrowns2004` dataset in metadat. 
This dataset contains results from 48 studies on the 
effectiveness of writing to learn interventions on academic achievement.
See `?metadat::dat.bangertdrowns2004` for details.

The dataset already contains computed `yi` and `vi` values, so no need for `escalc()`.

Fit a random effects model to these data.

```{r}
dat_g <- dat.bangertdrowns2004

mod_g <- rma(
  yi = yi ~ 1,
  vi = vi,
  slab = paste(id, author, year),
  method = "REML", test = "knha",
  data = dat_g
)
```

  1. Create a leave-1-out forest plot.
     Does leaving out any study substantially change the results?
     
```{r}
mod_loo <- leave1out(mod_g)

mod_loo

range(mod_loo$estimate)
range(sqrt(mod_loo$tau2))

with(mod_loo,
     forest(
       x = estimate,
       ci.lb = ci.lb, ci.ub = ci.ub,
       refline = predict(mod_g)$pred, slab = mod_g$slab,
       top = 0
     )
)
```

  - Leaving any one study out doesn't substantially change study results.
  
  2. Compute influence statistics for this model.
     Examine a table of values and plots of the values.
     Do any cases appear highly influential?
     
```{r}
mod_infl <- influence(mod_g)

mod_infl

plot(mod_infl)
```

  - Study 46 potentially sticks out, though even its Cook's D value is relatively small.
  
  3. Refit your model removing any influential cases.
     How do results change?
     
```{r}
mod_sensitivity <- update(mod_g, data = filter(dat_g, id != 46))
mod_sensitivity
```

  - Even leaving out this potential influential study does not change results.

  4. Examine the distribution of standardized residuals.
     Do these appear to be normally distributed?
          
```{r}
qqnorm(mod_g)
```

  - The standardized residuals look essentially normally distributed.

  5. Examine the distribution of estimated random effects.
     Do these appear to be normally distributed?
     
```{r}
qqnorm(ranef(mod_g)$pred)
qqline(ranef(mod_g)$pred)
```

  - The random effects appear to have somewhat wider tails than expected by a
    normal distribution, but given the number of studies, this deviation is not
    that severe.

Fit a mixed effects model to the full dataset including length of treatment as
a moderator. 
     
```{r}
dat_length <- drop_na(dat_g, length)

mod_length <- rma(
  yi = yi ~ 1 + length,
  vi = vi,
  slab = paste(author, year),
  method = "REML", test = "knha",
  data = dat_length
)
```

  6. Examine the distribution of standardized residuals and 
     estimated random effects for this moderated model.
     Do these appear to be normally distributed?
     
```{r}
qqnorm(mod_length)

qqnorm(ranef(mod_length)$pred)
qqline(ranef(mod_length)$pred)
```

 - Yes, both of these appear to be normally distributed. Accounting for treatment
   length removes the overly side tails from the random effect estimates distribution.

  7. Does length appear to be linearly related to the outcome measure?
     Are the residuals and random effects homogeneous/homoskedastic?
     
```{r}
# Linearity
data.frame(
  fitted = fitted(mod_length),
  residuals = residuals(mod_length, type = "rstudent")
) |>
ggplot() +
  aes(x = fitted, y = residuals) +
  labs(
    x = "Predicted values", y = "Standardized residuals"
  ) +
  theme_bw() +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth()

# Homogeneity of residuals
data.frame(
  fitted = fitted(mod_length),
  sqrt_resid = sqrt(abs(residuals(mod_length, type = "rstudent")))
) |>
  ggplot() +
  aes(x = fitted, y = sqrt_resid) +
  labs(
    x = "Predicted values", y = "sqrt( | Std. residuals | )"
  ) +
  theme_bw() +
  geom_point() +
  geom_smooth()

# Homogeneity of random effects
data.frame(
  fitted = fitted(mod_length),
  sqrt_ranef = sqrt(abs(ranef(mod_length)$pred))
) |>
  ggplot() +
  aes(x = fitted, y = sqrt_ranef) +
  labs(
    x = "Predicted values", y = "sqrt( | Random effects | )"
  ) +
  theme_bw() +
  geom_point() +
  geom_smooth()
```

  - Treatment length appears to be essentially linearly related to treatment effects.
  - Both the residuals and random effects appear to have somewhat lower variance
    when length is longer. It may be worthwhile to fit a location-scale model,
    though the number of studies with long treatments is small, so this result
    might must be noise.

  8. Compute influence statistics for this model.
     Do any of the meta-regression coefficients change substantially when
     one case is removed?
     
```{r}
mod_infl <- influence(mod_length)

mod_infl$dfbs

plot(mod_infl)
```

  - Study 46 (Willey 1988, study 1) again has somewhat large Cook's D. 
    However, its DFBETAS values for the intercept and slope for treatment length
    are not large (< 1).

  9. Remove any influential cases and re-estimate the mixed effects model
     How do results change?
     
```{r}
mod_sensitivity <- update(mod_length, data = filter(dat_length, id != 46))
mod_sensitivity
```

  - Even removing this one potential outlier does not substantially alter results
    for either the intercept or the length slope.
